Emissions and Perceptions of Climate Change

### 2022-12-14 Code/FINAL_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
done with intro!
Finished Final CAP DAP.

Statistical analysis will consist of the aforementioned three sections.

@@

Pulling in the corrected Emissions data. For the raw data see the raw data folder on github. Please also see the metadata txt files per each data set in the metadata folder.

@@

@@

Running a _______ statistical test will let us compare the rate of decrease in emissions of both these countries.

I talked to Robin Lock about this. We're going to do a paired t-test where each year has a pair of values (US, Denmark) and the values will be the RATE of change for that year, rather than the GHG emissions themselves.

Here's some info: https://www.statology.org/paired-samples-t-test-r/

```{r}
emissions$rate<-NA
denmark_rate<- emissions %>% filter(country=="Denmark")
den_rate<- diff(denmark_rate$emissions)
denmark_rate$rate[2:30]<-den_rate
denmark_rate<-denmark_rate[2:30,]

@@

```{r}
ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  geom_line() +
  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, and group tells R to group the points by country so geom_line will draw lines connecting the data by country.
```

```{r}
t.test(rate~country, data=emissions_rate, paired=TRUE)
#complete.cases means just use the ones without NAs
```

```{r}
t.test(rate~country, data=emissions_rate_2010, paired=TRUE)
```

We can also compare the moving averages and compare the rates, even forecasting to future dates (like what will emissions be in 2050 or 2100 based on the current time series). This comparision will be visual for the moving averages and just thoughtful for comparing the forecast numbers and will not involve a specific statistical test.

#### Running the Test

First determine the minimum date
```{r}
min_date=min(emissions$year)
max_date=max(emissions$year)
```
I'm not sure you'll need the above code chunk

Turning our data into a time series

```{r}
emissions_ts<-ts(emissions$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
```

Lets do time series for the indiviual countries
```{r}
emissions_Denmark<- emissions %>% filter(country=="Denmark")
emissions_US<-emissions %>% filter(country=="United States")
emissions_Denmark_ts<-ts(emissions_Denmark$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
emissions_US_ts<-ts(emissions_US$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
```

Viewing the time series of each individual country to ensure its correct.
```{r, results='hide'}
both_ts<-cbind(emissions_Denmark_ts, emissions_US_ts)
head(both_ts) # ensuring it worked
plot(both_ts, xlab="year", ylab="emissions")
```

If you wish to plot time series using ggplot, you can use autoplot(timeseriesname) and save it as an object. The result will be a ggplot object that you can manipulate. Example:

```{r}
demo_plot<-autoplot(emissions_US_ts)
demo_plot+
  labs(x = "Year", y = "Greenhouse gas emissions")+
  theme_bw()
```

And if you want to plot two time series on the same graph using ggplot, you use cbind to bind your two time series objects into the same new data frame and then autoplot that data frame:

```{r}
both<-cbind(emissions_Denmark_ts, emissions_US_ts)
autoplot(both)
```

Here is our data smoothed over so we can see the trends without added noise. The bottom one I presume is denmark. We can see that it is decreasing more.

Making a moving average:
```{r}
Ma_emissions<- forecast::ma(emissions_ts, order=3, centre = TRUE)
Ma_emissions_Denmark<- forecast:: ma(emissions_Denmark_ts, order=3, centre=TRUE)
Ma_emissions_US<- forecast::ma(emissions_US_ts, order=3, centre=TRUE)

Plotting both the moving average and the time series.

```{r, collapse=TRUE}
plot(emissions_ts)
lines(Ma_emissions, col="maroon", lwd=3)

plot(emissions_Denmark_ts)
lines(Ma_emissions_Denmark, col="orange", lwd=3)

plot(emissions_US_ts)
lines(Ma_emissions_US, col="blue")
```
Plotting both with autoplot/ggplot:

```{r}
USA<-cbind(emissions_US_ts, Ma_emissions_US)
autoplot(USA)
```

The red line is the data but smoothed, so we can see the overall trend of the emissions data combined. the other two are them separated, notice how similar Denmark's is to the combined data. Interesting.

***PUT RUNNING THE TIME SERIES HERE***
This will really be the paired t-test

### Perceptions of Climate Change

@@

Lets check with the 'beliefs question' which is the first question we will analyzed.

```{r}
#Parsing out just the survey answers regarding climate change beliefs
beliefs<- survey %>% filter(survey$question=="Ques__Climate_beliefs") 
```

@@

This means we can fairly compare Danes and American's perceptions of climate change.

I talked with Prof Shuckers about these data. He thinks we should do the back-calculating to change percents to numbers and then do a chi-square to compare answers between Danes and Americans. You will need the data on number of respondants for each question to do that.

???______
1101 people answered in Denmark
3096 answered in United States

```{r}
beliefs <- beliefs %>% group_by(Country)

dumb_way<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
beliefs<- cbind(beliefs, dumb_way)
beliefs<-beliefs %>% rename(num_people=...5)

beliefs<- beliefs %>% group_by(Country, response) %>%  mutate(total=(Number/100)*num_people)

beliefs$total<- round(beliefs$total, digits=0)

test<- beliefs %>% select(response, Country, total)
```

???? The chi-squared below did not work!!!
Chi-squared
```{r}
beliefs_matrix<-xtabs(total~Country + response, data=test)
beliefs_matrix
beliefs_df<- as.data.frame(beliefs_matrix)
beliefs_df<- beliefs_df %>% filter(Freq>0)
beliefs_matrix_real<- xtabs(Freq~Country+response, data=beliefs_df)
beliefs_matrix_real
chisq.test(beliefs_matrix)

```
???___

```{r}
vector_Den<- c(585,57,419,19,16,1)
vector_US<-c(1295,473,1073,86,161,7)
matrix<- as.matrix(rbind(vector_Den, vector_US))
chisq.test(matrix)

@@

Figure 1. What Danes and Americans believe caused climate change. Responses are in percent of the total surveyed.

More Americans believe that climate change is not happening or that it is caused by natural changes than Danes. More Danes believe it is caused by human activities. However, more Danes than Americans think it is cause equally by both. Lets look at a pie chart as well.

If you want to test these questions with a statistical test, we will need a combination of chi-square tests. Let's talk.

```{r}
piecolor<- c("orange","maroon","darkgreen","lightblue", "gray70","purple")

beliefs<- beliefs %>% arrange(desc(Number)) %>% mutate(Position=cumsum(Number/2)-0.5*Number)
legend<- c("caused equally","caused mostly by humans","caused mostly by natural changes","climate change isin't happening","other","refused") #see below
ggplot(beliefs, aes(x="", Percent, y= Number, fill=response))+  
  geom_bar(stat="identity", width=1, color="white")+
  coord_polar("y", start=0) + 
  facet_wrap(~Country)+
 #legend( 3, legend=legend)+, attempt to make the legends smaller, so the pie charts could be bigger and I could move the percentages.
  scale_fill_manual(values=piecolor )+
   theme_void()+# remove background, grid, numeric labels
  geom_label(aes (label = round(Number, digits = 1), x=1.7),color="white", size=4)  # add percentages

@@

Figure 2. Pie chart of what Danes and Americans believe caused climate change. Responses are in percent of the total surveyed.

Here's an article that talks more about data viz for likert type data. It has good arguments for why to avoid pie charts, and has a building series of charts ending with the author's (and my) favorite. Just worth considering:

https://nightingaledvs.com/visualizing-likert-scale-data-same-data-displayed-seven-different-ways/

The pie chart shows that most of the Danes surveyed believe climate change is human caused, whereas more Americans thought it was a mix of human and natural causes or that climate change is not happening.

We have multiple categorical predictor variables with the response variable being ratio data that bounded between 0 and 100%, so we will run a _________

Chi-square test after converting percents to numbers by multiplying the percents times the total number of repsondants to that question per country.

***INSERT ### Running the Statistical Test***

#### Level of worriedness about Climate Changed
In this survey question people were asked how worried they are about climate change. There are three options for 
```{r}
worry<- survey %>% filter(survey$question=="Ques__climate_worry") 
```

```{r}
ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity') + 
  labs(title="Level of Concern Over Climate Change", x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
```

Figure 3. Bar graph of the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.

32% of Americans are very worried about climate change, wherease oonly 21% of Danes are. But 52% of Danes are somewhat worried at only 36% are. To better visualize this lets try another graph.

```{r}
ggplot(worry, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
  theme_void()+ # remove background, grid, numeric labels
facet_wrap(~Country)
```

Figure 4. Pie chart of the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.

We can see that most Danes are somewhat worried, with the smallest percentage being not worried at all. Most Americans are also somewhat worried, though the percentage is smaller compared too Denmark (51% to 35.6%). More Americans however are very worried, and more Americans are not at all worried than Danes. Interesting. A statistical test is needed to see if Americans are less worried than Danes.

#### Lets do a statistical test to verify. 
Chi-square test. And think about a new data vis.

***### INSERT STATISTICAL TEST***

@@

Figure 5. Bar graph of the perceieved threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.

More Americans than Danes think that climate change is a very serious threat, but Danes think it is a somewhat serious threat than Americans. More Americans also think that climate change is not a threat at all than Danes. Lets view a pie chart.

```{r}
ggplot(threat, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.5, nudge_y = 2 ) + # add percentages
  theme_void()+ # remove background, grid, numeric labels
facet_wrap(~Country)
```

Figure 6. Pie char of the perceieved threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.

We can see that most Danes, %53 think climate change is a somewhat serious threat, wherease most Americans, %40.6, think it is a very serioous threat. Only 17.5% of Danes think that it is not a threat at all, compared to the 22.3% of Americans that think that. Lets test with a statistical test.

#### Lets do a statistical test to verify.

***###INSERT STATISTICAL TEST ***

### More or less Fossil Fuels

@@

Lets check with the 'beilefs question' which is the first we will anaylzed.

@@

More Americans think there should be much more or or somewhat more fossil fuels used than Danes. More Danes thinks threre should be somewhat less or much less fossil fuels used. This appears to clearly show that Danes think fossil fuel usage should be decreased, where at Americans think that less. Lets see in pie char form.

```{r}
ggplot(fuels, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
  theme_void()+ # remove background, grid, numeric labels
facet_wrap(~Country)
```

@@

The pie chart shows that most of the Danes surveyed, %53.6, believed much less fossil fuels should be used, with the second largest percentage of 25.7% believed somewhat less should be used. These percentages are larger than results in the United states, though the largest percentage of Americans also believed there should be much less fossil fuel usage at 40.5%, with the second most thinking somewhat less fossil fuel usage, 17.9%. More Americans believed there should be much more or somewhat more than Danes. Lets verify with a statistical test.

we will run another generalized linear model.

#### Running the Statistical Test

***### INSERT STATISTICAL TEST***

## Survey data Impacting Emissions

@@

## Limitations
The data used is from a survey conducted on Facebook, so all respondents are Facebook users. This adds a confounding variable, when utilizing this study.

>>>>

Statistical analysis will consist of the aforementioned two sections.

@@

Pulling in the corrected Emissions data. For the raw data see the raw data folder on github. Please also see the metadata txt files per each data set in the Data for the Code folder and the metadata folder.

@@

```
Figure 1. Per capita greenhouse gass emissions od Denmark and the U.S.from 1990 to 2019.

@@

Running a paired t test will let us compare the rate of decrease in emissions of both these countries.

### Running the Statistical Test

```{r}
t.test(emissions~country, data=emissions, paired=TRUE)
#like running a t test, but tell r the data is paired.
```
The p-value is very less than 2.2*10^-16, much less than 0.05, so we can reject our Null hypothesis. Denmark does have significantly less GHG emissions per capita than the United States.

## What about the Rate of Decrease?

Calculating rate

```{r}
emissions$rate<-NA
denmark_rate<- emissions %>% filter(country=="Denmark")
den_rate<- diff(denmark_rate$emissions) # subtracts n by n-1 to get rate.
denmark_rate$rate[2:30]<-den_rate
denmark_rate<-denmark_rate[2:30,]

@@

Lets look at the rate of Denmark and the United States.
```{r}
ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  geom_line() +
  geom_point(aes(colour=country))+ theme_bw() #colour colors by country, and group tells R to group the points by country so geom_line will draw lines connecting the data by country.
```
Figure 2. The rate of per capita greenhouse gas emissions for Denmark and the U.S. from 1990 to 2019.

There us alot of added noise, making it difficult to see overal trends. Lets view the rate data smoothed over.

```{r}
ggplot(emissions_rate, aes(year, rate, colour= country, group=country)) +  
  geom_smooth() + theme_bw() #colour colors by country, and group tells R to 
#group the points by country so geom_line will draw lines connecting the 
#data by country.

```
Figure 3. The smoothed over rate of per capita greenhouse gas emissions for Denmark and the U.S. from 1990 to 2019.

Denmark's rate is more variable over time, but it stays below 0, after 1995, wherease the US is not consistently below zero until 2000. It also goes above zero at 1990. Therefore, Denmark started decreasing its emissions five years prior to the U.S. and the U.S.'s emissions might be increasing again in 2019. Given Denmark has a lower rate from 2010-2019, it may have a significantly lower rate than the U.S. Lets run a statistical test to verify.

### Running the Statistical Test
Lets run a paired t-test to see if this is significant.
```{r}
t.test(rate~country, data=emissions_rate, paired=TRUE)
#complete.cases means just use the ones without NAs
```

Looking at the entire range from 1990-2019, the p value is 0.7668, so we must accept the Null hypothesiss. There is no significant difference in rate between Denmark and the U.S. Lets look at 2010-2019 given the differences in the funcitons listed above.

```{r, echo=FALSE}
# lets just included 210-2019 data.
emissions_rate_2010<- emissions_rate %>% 
  filter(emissions_rate$year>="2010-01-01")
```

```{r}
t.test(rate~country, data=emissions_rate_2010, paired=TRUE)
```
The p value is smaller at 0.083, however it is still not signifciant. So, we must accept the Null hypothesis. There is no significant difference in rate between Denmark and the United States.

### Conclusion for Section One

The United States emits significantly more greenhouse gases per capita than Denmark, but there was no significant difference in how fast both countries decreased their emissions from 1990-2019.

Now lets look at some examples that may explain why Denmark emits less greenhouse gases per capita.

## Perceptions of Climate Change

@@

Lets check with the 'beliefs question' which is the first question we will analyze to ensure Denmark and the U.S. are comparable despite size differences.

```{r}
#Parsing out just the survey answers regarding climate change beliefs
beliefs<- survey %>% filter(survey$question=="Ques__Climate_beliefs")

#Resetting the levels

beliefs$response<-factor(beliefs$response)
levels(beliefs$response)
```

@@

@@

Figure 4. What Danes and Americans believe caused climate change. Responses are in percent of the total surveyed.

More Americans believe that climate change is not happening or that it is caused by natural changes than Danes. More Danes believe it is caused by human activities. However, more Danes than Americans think it is cause equally by both.

To get a better sense of the of which country answeres were more beneficial for combating climate change, lets look at a diverging bar graph. In this graph answers that increase our chances of combating climate change, in this case believing that climate change is caused by human activity, is made positive and answers that decrease our chances of combating climate change, in this case all other answers, are made negative. Due to the small percentage of people who refused, those will be removed from the diverging bar graph, so it is more visually appealing.

```{r}
beliefs_diverging <- beliefs %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Caused mostly by human activities"), Number, -Number)) 
#all the percentages in the response column that did not answer "caused mostly by human activities"   will have a - put in front. 
```

```{r}
beliefs_diverging_Order<-beliefs_diverging %>% mutate(response=fct_relevel(response, "Caused about equally by human activities and natural changes", "Other", "Caused mostly by natural changes in the environment", "None of the above because climate change isn’t happening", "Caused mostly by human activities"), response=fct_rev(response))

beliefs_diverging_Order
```

```{r}

#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
beliefs_diverging_Order<- beliefs_diverging_Order %>% mutate(percent=paste0(round(abs(Number), digits=1), "%"))

beliefs_divering<-ggplot(beliefs_diverging_Order, aes(x=Country, y=Number, fill= response)) + 
  geom_col()+ #also makes a bar graph
  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
  coord_flip()+
  scale_x_discrete()+

  #reorganizes the labels and color them
  scale_fill_manual(breaks=c("None of the above because climate change isn’t happening","Caused mostly by natural changes in the environment", "Other", "Caused about equally by human activities and natural changes","Caused mostly by human activities"), values=c(
    "None of the above because climate change isn’t happening"="darkorange4",
    "Caused mostly by natural changes in the environment" = "darkorange3",
    "Other" = "darkorange", 
    "Caused about equally by human activities and natural changes"= "goldenrod1",
    "Caused mostly by human activities" = "deepskyblue"
  ))+

  labs(title="What do you Think Caused Climate Change", x=NULL, fill=NULL)+
 theme_minimal()+
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
panel.grid=element_blank(),
legend.position="top")

beliefs_divering

ggsave("beliefs_diverging.png")

```
Figure 4. Diverging Bar graph of what Danes and Americans believe caused climate change. Responses are in percent of the total surveyed.Darker orange is a more anti climate change answer, wherease blue is a climate activist answer.

For a easier to read figure please see the image saved in the code folder.

The diverging bar graph shows that most of the Danes surveyed believe climate change is human caused, whereas more Americans thought it was a mix of human and natural causes or that climate change is not happening.

### Running the Statistical Test

We can put our data into two categories: those that believe climate change is human caused, and all other answers. This way we can run a chi-squared test and see if if it is significant.

first must convert percents to numbers by multiplying the percents times the total number of respondents to that question per country.

```{r, echo=FALSE}
respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
#1101 Danes answered the survey, 3096 Americans did.
beliefs$num_people<- respondants #put that vector into here

beliefs<- beliefs %>% group_by(Country, response) %>% 
  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage

beliefs$total<- round(beliefs$total, digits=0) #round to whole number

beliefs_numPeople<- beliefs%>% select(response, Country, total)
```

```{r}

beliefs_numPeople$grouped_answer<- NA
good<- which( beliefs_numPeople$response== "Caused mostly by human activities")
bad<-which(beliefs_numPeople$response=="Caused mostly by natural changes in the environment" |beliefs_numPeople$response=="Caused about equally by human activities and natural changes"
           |beliefs_numPeople$response=="Other"
           |beliefs_numPeople$response=="None of the above because climate change isn’t happening")
  beliefs_numPeople$grouped_answer[good]<- "good"
  beliefs_numPeople$grouped_answer[bad]<-"bad"

beliefs_numPeople<- beliefs_numPeople%>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))

@@

```{r}
beliefs.mat<-xtabs(grouped_number~Country + grouped_answer, data=beliefs_numPeople)
beliefs.mat
chisq.test(beliefs.mat)
```
The p values is 3.829*10^-11, which is much less than 95%, so I can reject the Null hypothesis. Significantly more Danes think climate change is caused by humans verses other sources than Americans.

## Level of worriedness about Climate Changed

In this survey question people were asked how worried they are about climate change. There are three options, very worried, somewhat worried, not very worried, not at all worried, and refused.
```{r}
worry<- survey %>% filter(survey$question=="Ques__climate_worry") 
```

```{r}
ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
  labs(title="Level of Concern Over Climate Change", x="Response to Survey Question", y="Percent") + theme_bw() + coord_flip()
```

Figure 5. Bar graph of the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.

32% of Americans are very worried about climate change, whereas only 21% of Danes are. But 52% of Danes are somewhat worried at only 36% are. To better visualize this lets look at a diverging bar graph.

Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.
```{r}
worry_diverging <- worry %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Very worried", "Somewhat worried"), Number, -Number)) 
#all the percentages in the response column that did not answer "caused mostly by human activities"   will have a - put in front. 
```

```{r}
worry_diverging<-worry_diverging %>% mutate(response=fct_relevel(response, "Not at all worried","Not very worried", "Somewhat worried", "Very worried"), response=fct_rev(response))

```{r}

#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
worry_diverging<- worry_diverging %>% mutate(percent=paste0(round(abs(Number), digits=1), "%"))

ggplot(worry_diverging, aes(x=Country, y=Number, fill= response)) + 
  geom_col()+ #also makes a bar graph
  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
  coord_flip()+
  scale_x_discrete()+

  #reorganizes the labels and color them
  scale_fill_manual( breaks=c("Not at all worried","Not very worried", "Somewhat worried", "Very worried"),
      values=c(
    "Not at all worried"="darkorange4",
    "Not very worried" = "darkorange3",
    "Somewhat worried"= "deepskyblue",
    "Very worried" = "deepskyblue3"
  ))+

  labs(title="How Much Does Climate Change Worry You", x=NULL, fill=NULL)+
 theme_minimal()+
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
panel.grid=element_blank(),
legend.position="top")

Figure 6. Diverging Bar Graph the Level of Concern over climate change of both Danes and Americans. Response is in percentage of the total surveyed.

We can see that most Danes are somewhat worried, with the smallest percentage being not worried at all. Most Americans are also somewhat worried, though the percentage is smaller compared too Denmark (51% to 35.6%). More Americans however are very worried, and more Americans are not at all worried than Danes. Based on the diverging bar graph, there is a slight difference between countries, suggesting Danes are more worried than Americans. A statistical test is needed to see if this is a significant difference.

### Lets do a statistical test to verify. 
Chi-square test.

Very and somewhat worried will be grouped together and not at all worried and not very worried will be grouped together.

first must convert percents to numbers by multiplying the percents times the total number of respondents to that question per country.

```{r, echo=FALSE}
respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
#1101 Danes answered the survey, 3096 Americans did.
worry$num_people<- respondants #put that vector into here

worry<- worry %>% group_by(Country, response) %>% 
  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage

worry$total<- round(worry$total, digits=0) #round to whole number

worry_numPeople<- worry%>% select(response, Country, total)
```

```{r}

worry_numPeople<- worry_numPeople %>% filter(response!="Refused")
worry_numPeople$grouped_answer<- NA
not<- which(worry_numPeople$response== "Not at all worried"|worry_numPeople$response== "Not very worried")
worried<- which( worry_numPeople$response== "Very worried" |worry_numPeople$response== "Somewhat worried")

  worry_numPeople$grouped_answer[worried]<- "worried"
  worry_numPeople$grouped_answer[not]<- "not"

worry_numPeople<- worry_numPeople%>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))

```{r}
worry.mat<-xtabs(grouped_number~Country+grouped_answer, data=worry_numPeople)
worry.mat
chisq.test(worry.mat)
```
the p-value is 0.01258, which is less than 0.05. So, we can reject the Null hypothesis. Danes are more worried about climate change than Americans.

@@

Figure 7. Bar graph of the perceieved threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.

More Americans than Danes think that climate change is a very serious threat, but more Danes think it is a somewhat serious threat than Americans. More Americans also think that climate change is not a threat at all than Danes. Lets view a diverging bar graph.

Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.
```{r}
threat_diverging <- threat %>% filter(response!="Refused") %>% mutate(Number=if_else(response %in% c("Very serious threat", "Somewhat serious threat"), Number, -Number))

Reorganize the bars.
```{r}
threat_diverging<-threat_diverging %>% mutate(response=fct_relevel(response, "Don't know", "Not a threat at all", "Somewhat serious threat", "Very serious threat"), response=fct_rev(response))

```{r}
#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
threat_diverging<- threat_diverging %>% mutate(percent=paste0(round(abs(Number), digits=1), "%"))

ggplot(threat_diverging, aes(x=Country, y=Number, fill= response)) + 
  geom_col()+ #also makes a bar graph
  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
  coord_flip()+
  scale_x_discrete()+

  #reorganizes the labels and color them
  scale_fill_manual( breaks=c("Not a threat at all","Don't know", "Somewhat serious threat", "Very serious threat"),
      values=c(
    "Not a threat at all"= "darkorange3",
    "Don't know" = "goldenrod3",
    "Somewhat serious threat"= "deepskyblue",
    "Very serious threat" = "deepskyblue3"
  ))+

  labs(title="In the Next 20 Years How Much of a Threat is Climate Change?", x=NULL, fill=NULL)+
 theme_minimal()+
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
panel.grid=element_blank(),
legend.position="top")

Figure 8. Diverging bar graph of the perceived threat of climate change of Danes and Americans. Response is in percentage of the total surveyed.

We can see that most Danes, 53%, think climate change is a somewhat serious threat, wherease most Americans, 40.6%, think it is a very serious threat. Only 17.5% of Danes think that it is not a threat at all, compared to the 22.3% of Americans that think that. There is a small shift in Dane's bars leaning right, suggesting they think climate change is a more serious threat than Americans. Lets see if it is significant with a statistical test.

### Lets do a statistical test to verify.

first must convert percents to numbers by multiplying the percents times the total number of respondents to that question per country.

```{r, echo=FALSE}
respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
#1101 Danes answered the survey, 3096 Americans did.
threat$num_people<- respondants #put that vector into here

threat<- threat %>% group_by(Country, response) %>% 
  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage

threat$total<- round(threat$total, digits=0) #round to whole number

threat_numPeople<- threat%>% select(response, Country, total)
```

```{r}

threat_numPeople<- threat_numPeople %>% filter(response!="Refused") 
threat_numPeople$grouped_answer<- NA
no<- which(threat_numPeople$response== "Not a threat at all" |threat_numPeople$response== "Don't know" )
yes<- which(threat_numPeople$response== "Somewhat serious threat" |threat_numPeople$response== "Very serious threat")

  threat_numPeople$grouped_answer[no]<- "no"
  threat_numPeople$grouped_answer[yes]<- "yes"

threat_numPeople<- threat_numPeople%>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))

```{r}
threat.mat<-xtabs(grouped_number~Country+grouped_answer, data=threat_numPeople)
threat.mat
chisq.test(threat.mat)
```
Pretending 'Don't know' is a negative response, there is not a significant difference in how much of a threat climate change is perceived as by Danes and Americans.

```{r, echo=FALSE}
respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
#1101 Danes answered the survey, 3096 Americans did.
threat$num_people<- respondants #put that vector into here

threat2<- threat %>% group_by(Country, response) %>% 
  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage

threat2$total<- round(threat2$total, digits=0) #round to whole number

threat2_numPeople<- threat2%>% select(response, Country, total)
```

```{r, echo=FALSE}

threat2_numPeople<- threat2_numPeople %>% filter(response!="Refused" & response!="Don't know") 
threat2_numPeople$grouped_answer<- NA
no<- which(threat2_numPeople$response== "Not a threat at all" )
yes<- which(threat2_numPeople$response== "Somewhat serious threat" |threat2_numPeople$response== "Very serious threat")

  threat2_numPeople$grouped_answer[no]<- "no"
  threat2_numPeople$grouped_answer[yes]<- "yes"

threat2_numPeople<- threat2_numPeople%>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))

```{r, echo=FALSE}
threat2.mat<-xtabs(grouped_number~Country+grouped_answer, data=threat2_numPeople)
threat2.mat
chisq.test(threat2.mat)
```
Ignoring the Don't knows yields a p value of 0.005, so I can reject the Null hypothesis. Significantly more Danes think that within the next 20 years climate change is a serious threat.

## More or less Fossil Fuels

@@

@@

More Americans think there should be much more or or somewhat more fossil fuels used than Danes. More Danes thinks threre should be somewhat less or much less fossil fuels used. This appears to clearly show that Danes think fossil fuel usage should be decreased, where at Americans think that less. Lets see in diverging bar graph.

Given the small number of people who answered, refused, it will be left out for clairty of the diverging bar graph.

Again the option "Don't know" is present. This also has the option of same amount as today. According to the IPCC humans must decrease their emissions to stay well below 2 degrees of warming and avoid the most severe impacts of climate change, therefore, keeping it the same will go on the negative side of the diverging bar graph. Not many people chose "Don't know," plus it is difficult to predict, so I believe it is fair to leave it out.
```{r}
fuels_diverging <- fuels %>% filter(response!="Refused" & response!= "Don’t know") %>% mutate(Number=if_else(response %in% c("Much less", "Somewhat less"), Number, -Number))

Reorganize the bars.
```{r}
fuels_diverging<-fuels_diverging %>% mutate(response=fct_relevel(response,"Much more","Somewhat more", "Same amount as today", "Somewhat less", "Much less"), response=fct_rev(response))

```{r}
#first line below takes the Number column rounds it up, takes the absolute value of it and adds a % for the labels on the graph.
fuels_diverging<- fuels_diverging %>% mutate(percent=paste0(round(abs(Number), digits=0), "%"))
#rounded it to 0 this time due to serious overlap in the % labels on the graph

ggplot(fuels_diverging, aes(x=Country, y=Number, fill= response)) + 
  geom_col()+ #also makes a bar graph
  geom_text(aes(label=percent), position=position_stack(vjust = 0.5), color="white", fontface="bold")+ #stack stacks them, fontface bolds  the numbers in the bargraphs 
  coord_flip()+
  scale_x_discrete()+

  #reorganizes the labels and color them
  scale_fill_manual( breaks=c("Much more", "Somewhat more", "Same amount as today", "Somewhat less", "Much less"),
      values=c(
    "Much more"= "darkorange4",
    "Somewhat more"= "darkorange3",
    "Same amount as today" = "darkorange2",
    "Somewhat less"= "deepskyblue",
    "Much less" = "deepskyblue3"
  ))+

  labs(title="How Much Fossil Fuels Should Your Country Use?", x=NULL, fill=NULL)+
 theme_minimal()+
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(),
panel.grid=element_blank(),
legend.position="top")

@@

The diverging bar graph shows that most of the Danes surveyed, 54%, believed much less fossil fuels should be used, with the second largest percentage of 26% believed somewhat less should be used. These percentages are larger than results in the United states, though the largest percentage of Americans also believed there should be much less fossil fuel usage at 40%, with the second most thinking somewhat less fossil fuel usage, 18%. More Americans believed there should be much more or somewhat more fossil fuel usage than Danes. There is also a visible difference in the overall diverging bar graphs, sugesting that more Danes believe their country should use less fossil fuels, than Americans believed. Lets verify with a statistical test.

we will run chi-squared test.

### Running the Statistical Test

```{r, echo=FALSE}
respondants<- c(1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096,1101,3096)
#1101 Danes answered the survey, 3096 Americans did.
fuels$num_people<- respondants #put that vector into here

fuels<- fuels %>% group_by(Country, response) %>% 
  mutate(total=(Number/100)*num_people) #calculating rate where Number=percentage

fuels$total<- round(fuels$total, digits=0) #round to whole number

fuels_numPeople<- fuels%>% select(response, Country, total)
```

```{r, echo=FALSE}

fuels_numPeople<- fuels_numPeople %>% filter(response!="Refused" & response!="Don't know") 
fuels_numPeople$grouped_answer<- NA
more<- which(fuels_numPeople$response== "Much more" | fuels_numPeople$response== "Somewhat more" | fuels_numPeople$response== "Same amount as today" )
less<- which(fuels_numPeople$response== "Much less" |fuels_numPeople$response== "Somewhat less")

  fuels_numPeople$grouped_answer[more]<- "more"
  fuels_numPeople$grouped_answer[less]<- "less"

fuels_numPeople<- fuels_numPeople%>% group_by(Country, grouped_answer) %>% summarise(grouped_number=sum(total))

```{r, echo=FALSE}
fuels.mat<-xtabs(grouped_number~Country+grouped_answer, data=fuels_numPeople)
fuels.mat
chisq.test(fuels.mat)
```
The p-value is 2.2*10^-16, much less than 0.05, meaning we can reject the null hypothesis. Significantly more Danes believe their country should use less fossil fuels, than Americans.

@@

Denmark emits significantly less per capita greenhouse emissions than the United states. My results also showed that significantly more Danes believed climate change was human caused, were more worried aobut climate change, and think their country should decrease fossil fuel usage than Americans did. Overall, Danes' answers to the survey questions were significantly more planet friendly than Americans. Danish citizen's perceptions of climate change could be way Denmark emmits so much less greenhouse gas emissions per person than the United States.

It was difficult to asses the to what degree in the next 20 years is climate change a threat question, due to the "don't know" option. With it, there was no significant difference in responses between the countries. If I do not include it Denmark thought climate change was a bigger threat than the United States. On one hand people who answered do not know probably don't partake in climate change related activities such as voting for policies, making their decisions neutral. On the other hand, it could mean the person would lean against climate policies, not really considering climate change a pertinent issue to them personally. It is difficult to determine how someone who answered don't know would react, so not including it seems reasonable. Furthermore, given Denmark answered significantly more favorably in the three other questions, a significant p value when removing the "don't know" responses is not surprising. However, it is interesting that more Danes answered did not know to this question than Americans.

## Limitations
